---
title: "RateMyProf"
output: html_document
date: "2024-11-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing Dependecies
#hello
```{r, echo = FALSE}
library(tidyverse)
```

Uploading the CSV for Rate My Professor Data and Daily Nexus Data

```{r, echo = FALSE}
rmpdata <- read_csv("/Users/laurenceliao/Downloads/teacher_ratings.csv")
nrow(rmpdata)
```
```{r}
coursegradesdata <- read_csv("/Users/laurenceliao/Downloads/courseGrades.csv")
nrow(coursegradesdata)
```


# Data Cleaning

- Changing the ratings column to remove the "ratings" at the end
- filter to only have those with ratings
- remove professor duplicates


```{r, echo = FALSE}
rmpdata <- rmpdata |> mutate(`Number of Ratings` = as.numeric(gsub("[^0-9]", "", `Number of Ratings`))) |> filter(`Number of Ratings` > 5)
nrow(rmpdata)
rmpdata <- rmpdata |> distinct(Name, .keep_all = TRUE)
nrow(rmpdata)
```

There were a total of 291 professors that had 0 ratings and 20 duplicate ratings. The number of professors we are now analyzing is 2804.

- Group grades so that each category is just A, B, C, D, F regardless of plus or minus

```{r}
coursegradesdata <- coursegradesdata |> mutate(totalA = A + Ap + Am,
                                               totalB = B + Bp + Bm,
                                               totalC = C + Cp + Cm, 
                                               totalD = D + Dp + Dm,
                                               totalF = F)
```

- Group quality ratings to 0-1 group, 1-2, etc.

```{r}
rmpdata <- rmpdata |> mutate(qualitygroup = 
  case_when(`Quality Rating` >= 0 & `Quality Rating` < 1 ~ "0-1",
            `Quality Rating` >= 1 & `Quality Rating` < 2 ~ "1-2",
            `Quality Rating` >= 2 & `Quality Rating` < 3 ~ "2-3",
            `Quality Rating` >= 3 & `Quality Rating` < 4 ~ "3-4",
            `Quality Rating` >= 4 & `Quality Rating` <= 5 ~ "4-5"),
 difficultygroup = case_when(Difficulty >= 0 & Difficulty < 1 ~ "0-1",
                             Difficulty >= 1 & Difficulty < 2 ~ "1-2",
                             Difficulty >= 2 & Difficulty < 3 ~ "2-3",
                             Difficulty >= 3 & Difficulty < 4 ~ "3-4",
                             Difficulty >= 4 & Difficulty <= 5 ~ "4-5"))
```


# Classifying Departments into 5 Main Subgroups: HSSA, Science, Engineering, Math/Stats, Economics/Business

```{r}

dept_mapping <- c(
  ANTH = "HSSA",
  ART = "HSSA",
  ARTHI = "HSSA",
  ARTHIW = "HSSA",
  ARTST = "HSSA",
  AS = "HSSA",
  ASTRO = "Science",
  ASTROW = "Science",
  BIOE = "Engineering",
  BIOL = "Science",
  BL = "HSSA",
  BMSE = "Science",
  C = "HSSA",
  CH = "HSSA",
  CHEM = "Science",
  CHIN = "HSSA",
  CLASS = "HSSA",
  CMPSC = "Engineering",
  CMPSCCS = "Engineering",
  CMPSCW = "Engineering",
  CMPTGCS = "Engineering",
  CNCSP = "HSSA",
  CNCSPW = "HSSA",
  COMM = "HSSA",
  DANCE = "HSSA",
  DANCEW = "HSSA",
  #drop DYNS = "Science",
  EACS = "HSSA",
  EARTH = "Science",
  EARTHW = "Science",
  ECE = "Engineering",
  ECON = "Economics/Business",
  ED = "HSSA",
  EDS = "Math/Stats",
  EEMB = "Science",
  ENGL = "HSSA",
  ENGR = "Engineering",
  ENV = "Science",
  ES = "HSSA",
  ESS = "HSSA",
  ESM = "Economics/Business",
  FAMST = "HSSA",
  FAMSTW = "HSSA",
  FEMST = "HSSA",
  FEMSTW = "HSSA",
  FLMST = "HSSA",
  FR = "HSSA",
  GEOG = "Science",
  GEOL = "Science",
  GER = "HSSA",
  GLOBL = "HSSA",
  GPS = "HSSA",
  GREEK = "HSSA",
  # drop GRAD = "HSSA",
  HEB = "HSSA",
  HIST = "HSSA",
  INT = "HSSA",
  ITAL = "HSSA",
  JAPAN = "HSSA",
  JAPANW = "HSSA",
  KOR = "HSSA",
  LAIS = "HSSA",
  LATIN = "HSSA",
  #LAW = "HSSA",
  LING = "HSSA",
  LIT = "HSSA",
  MARSC = "Science",
  MAT = "Science",
  MATH = "Math/Stats",
  MATRL = "Science",
  MCDB = "Science",
  ME = "Engineering",
  MES = "HSSA",
  MS = "HSSA",
  MUS = "HSSA",
  PHIL = "HSSA",
  PHYS = "Science",
  POL = "HSSA",
  PORT = "HSSA",
  PSTAT = "Math/Stats",
  PSTATW = "Math/Stats",
  PSY = "HSSA",
  RG = "HSSA",
  RUSS = "HSSA",
  SHS = "HSSA",
  SLAV = "HSSA",
  SOC = "HSSA",
  SPAN = "HSSA",
  THTR = "HSSA",
  TMP = "Economics/Business",
  "W&L" = "HSSA",
  WRIT = "HSSA"
)

coursegradesdata <- coursegradesdata %>%
  mutate(dep_group = dept_mapping[dept])
```

```{r}
dept_mapping_rmp <- c( 
  "Accounting" = "Economics/Business", 
  "African-American Studies" = "HSSA", 
  "Anthropology" = "HSSA", 
  "Architecture" = "HSSA", 
  "Art" = "HSSA", 
  "Art History" = "HSSA", 
  "Asian American Studies" = "HSSA", 
  "Biochemistry" = "Science", 
  "Biology" = "Science", 
  "Black Studies" = "HSSA", 
  "Business" = "Economics/Business", 
  "Chemical Engineering" = "Engineering", 
  "Chemistry" = "Science", 
  "Chicano Studies" = "HSSA", 
  "Classics" = "HSSA", 
  "Communication" = "HSSA", 
  "Computer Science" = "Engineering", 
  "East Asian Languages & Cultures" = "HSSA", 
  "Economics" = "Economics/Business", 
  "Education" = "HSSA", 
  "Electrical & Comp. Engineering" = "Engineering", 
  "Engineering" = "Engineering", 
  "English" = "HSSA", 
  "Environmental Science" = "Science", 
  "Environmental Science & Mgmt." = "Science", 
  "Environmental Studies" = "Science", 
  "Ethnic Studies" = "HSSA", 
  "Exercise & Sport Studies" = "HSSA", 
  "Feminist Studies" = "HSSA", 
  "Film" = "HSSA", 
  "Finance" = "Economics/Business", 
  "Fine Arts" = "HSSA", 
  "French" = "HSSA", 
  "French & Italian" = "HSSA", 
  "Geography" = "Science", 
  "Geology" = "Science", 
  "Germanic & Russian Studies" = "HSSA", 
  "Global & International Studies" = "HSSA", 
  "Global Studies" = "HSSA", 
  "Graphic Arts" = "HSSA", 
  "Health Science" = "Science", 
  "History" = "HSSA", 
  "Humanities" = "HSSA", 
  "Information Science" = "Engineering", 
  "International Studies" = "HSSA", 
  "Languages" = "HSSA", 
  #"Law" = "HSSA", 
  "Life Science" = "Science", 
  "Linguistics" = "HSSA", 
  "Literature" = "HSSA", 
  "Mathematics" = "Math/Stats", 
  "Mechanical Engineering" = "Engineering", 
  "Music" = "HSSA", 
  "Philosophy" = "HSSA", 
  "Physical Education" = "HSSA", 
  "Physics" = "Science", 
  "Political Science" = "HSSA", 
  "Psychology" = "HSSA", 
  "Religion" = "HSSA", 
  "Religious Studies" = "HSSA", 
  "Science" = "Science", 
  "Social Science" = "HSSA", 
  "Sociology" = "HSSA", 
  "Spanish" = "HSSA", 
  "Spanish & Portuguese" = "HSSA", 
  "Statistics" = "Math/Stats", 
  "Theater" = "HSSA", 
  "Theology" = "HSSA", 
  "Women's Studies" = "HSSA", 
  "Writing" = "HSSA"
)

rmpdata <- rmpdata %>%
  mutate(dep_group = dept_mapping_rmp[Department])
```



```{r}
coursegradesdata <- coursegradesdata[!is.na(coursegradesdata$dep_group), ]
rmpdata <- rmpdata[!is.na(rmpdata$dep_group), ]
```


# Merging with new groupings of departments
```{r}
rmpdatatest <- rmpdata |> mutate(nameTransformed = str_to_upper(str_c(word(Name, -1), " ", substr(word(Name, 1), 1, 1))))
coursegradesdatatest <- coursegradesdata |> 
  mutate(nameTransformed = if_else(str_count(instructor, "\\s+") == 2, str_remove(instructor, " \\w+$"), instructor))

# Removing extra dep_group from rmp since it already is there for course grades. Would this run into issues in the future?
rmpdatatest <- rmpdatatest |> select(-dep_group)
finaldata <- inner_join(rmpdatatest, coursegradesdatatest, by = "nameTransformed") |> select(-instructor, -nameTransformed, -Department)

```

# HSSA Department Group 

```{r}
# Filter such that only letter graded classes show up
HSSA_data <- finaldata |> filter(dep_group == "HSSA" & nLetterStudents > 0)
grouped_HSSA_data <- HSSA_data |> group_by(Name) |> 
  summarize(A = sum(totalA),
            B = sum(totalB), 
            C = sum(totalC),
            D = sum(totalD),
            F = sum(totalF)) |>
  ungroup() |> mutate(totalGrades = A + B + C + D + F)

# These few lines of code are to join the quality and difficulty ratings of each professor as found in rate my professor
HSSA_data_list <- HSSA_data |> distinct(Name, .keep_all = TRUE) |> arrange(Name) |> select(Name, qualitygroup, difficultygroup, dep_group) 
HSSA_final_data <- grouped_HSSA_data |> inner_join(HSSA_data_list, by = "Name")
```

# Science Department Group
```{r}
Science_data <- finaldata |> filter(dep_group == "Science" & nLetterStudents > 0)
grouped_Science_data <- Science_data |> group_by(Name) |> 
  summarize(A = sum(totalA),
            B = sum(totalB), 
            C = sum(totalC),
            D = sum(totalD),
            F = sum(totalF)) |>
  ungroup() |> mutate(totalGrades = A + B + C + D + F)


Science_data_list <- Science_data |> distinct(Name, .keep_all = TRUE) |> arrange(Name) |> select(Name, qualitygroup, difficultygroup, dep_group) 
Science_final_data <- grouped_Science_data |> inner_join(Science_data_list, by = "Name")
```

# Engineering Department Group 

```{r}
Engineering_data <- finaldata |> filter(dep_group == "Engineering" & nLetterStudents > 0)
grouped_Engineering_data <- Engineering_data |> group_by(Name) |> 
  summarize(A = sum(totalA),
            B = sum(totalB), 
            C = sum(totalC),
            D = sum(totalD),
            F = sum(totalF)) |>
  ungroup() |> mutate(totalGrades = A + B + C + D + F)

Engineering_data_list <- Engineering_data |> distinct(Name, .keep_all = TRUE) |> arrange(Name) |> select(Name, qualitygroup, difficultygroup, dep_group) 
Engineering_final_data <- grouped_Engineering_data |> inner_join(Engineering_data_list, by = "Name")
Engineering_final_data
```

# Math/Stats Department Group
```{r}
Math_Stats_data <- finaldata |> filter(dep_group == "Math/Stats" & nLetterStudents > 0)
grouped_Math_Stats_data <- Math_Stats_data |> group_by(Name) |> 
  summarize(A = sum(totalA),
            B = sum(totalB), 
            C = sum(totalC),
            D = sum(totalD),
            F = sum(totalF)) |>
  ungroup() |> mutate(totalGrades = A + B + C + D + F)

Math_Stats_data_list <- Math_Stats_data |> distinct(Name, .keep_all = TRUE) |> arrange(Name) |> select(Name, qualitygroup, difficultygroup, dep_group) 
Math_Stats_final_data <- grouped_Math_Stats_data |> inner_join(Math_Stats_data_list, by = "Name")
```
# Economics/Business Department Data

```{r}
Economics_Business_data <- finaldata |> filter(dep_group == "Economics/Business" & nLetterStudents > 0)
grouped_Economics_Business_data <- Economics_Business_data |> group_by(Name) |> 
  summarize(A = sum(totalA),
            B = sum(totalB), 
            C = sum(totalC),
            D = sum(totalD),
            F = sum(totalF)) |>
  ungroup() |> mutate(totalGrades = A + B + C + D + F)

Economics_Business_data_list <- Economics_Business_data |> distinct(Name, .keep_all = TRUE) |> arrange(Name) |> select(Name, qualitygroup, difficultygroup, dep_group) 
Economics_Business_final_data <- grouped_Economics_Business_data |> inner_join(Economics_Business_data_list, by = "Name")

Economics_Business_final_data
```


