---
title: "RateMyProf"
output: html_document
date: "2024-11-14"
---

# Installing Dependecies
```{r, echo = FALSE}
library(tidyverse)
```

# Reading in Data
```{r, echo = FALSE, warning=FALSE, message=FALSE}
rmpdata <- read_csv("teacher_ratings.csv")
gradesdata <- read_csv("courseGrades.csv")

nrow(gradesdata)
```

# Data Cleaning
Here, we need to:
- Changing the values in the ratings column to remove the "ratings" string at the end
- Filter to only have those with 5 or more ratings
- Remove duplicate entries

```{r, echo = FALSE}
rmp_clean <- rmpdata |> 
  mutate(`Number of Ratings` = as.numeric(gsub("[^0-9]", "", `Number of Ratings`))) |> 
  filter(`Number of Ratings` > 5) |> 
  distinct(Name, .keep_all = TRUE)

head(rmp_clean)
```
There were a total of 291 professors that had 0 ratings and 20 duplicate ratings. The number of professors we are now analyzing is 2804.

Next, we will mutate columns which are ranges of `Quality Rating` and `Difficulty Rating`.
```{r}
rmp_clean <- rmp_clean |>
  mutate(qualitygroup = 
           case_when(`Quality Rating` >= 0 & `Quality Rating` < 1 ~ "0-1",
                     `Quality Rating` >= 1 & `Quality Rating` < 2 ~ "1-2",
                     `Quality Rating` >= 2 & `Quality Rating` < 3 ~ "2-3",
                     `Quality Rating` >= 3 & `Quality Rating` < 4 ~ "3-4",
                    `Quality Rating` >= 4 & `Quality Rating` <= 5 ~ "4-5"),
         difficultygroup = 
           case_when(Difficulty >= 0 & Difficulty < 1 ~ "0-1",
                     Difficulty >= 1 & Difficulty < 2 ~ "1-2",
                     Difficulty >= 2 & Difficulty < 3 ~ "2-3",
                     Difficulty >= 3 & Difficulty < 4 ~ "3-4",
                     Difficulty >= 4 & Difficulty <= 5 ~ "4-5"))
```

Similarly, we  aggregate grades so that each category is just A, B, C, D, F regardless of plus or minus
```{r}
grades_clean <- gradesdata |> 
  mutate(totalA = A + Ap + Am,
         totalB = B + Bp + Bm,
         totalC = C + Cp + Cm, 
         totalD = D + Dp + Dm,
         totalF = F)
```

# Classifying Departments
```{r}
dept_mapping <- c(
  ANTH = "HSSA",
  ART = "HSSA",
  ARTHI = "HSSA",
  ARTHIW = "HSSA",
  ARTST = "HSSA",
  AS = "HSSA",
  ASTRO = "Science",
  ASTROW = "Science",
  BIOE = "Engineering",
  BIOL = "Science",
  BL = "HSSA",
  BMSE = "Science",
  C = "HSSA",
  CH = "HSSA",
  CHEM = "Science",
  CHIN = "HSSA",
  CLASS = "HSSA",
  CMPSC = "Engineering",
  CMPSCCS = "Engineering",
  CMPSCW = "Engineering",
  CMPTGCS = "Engineering",
  CNCSP = "HSSA",
  CNCSPW = "HSSA",
  COMM = "HSSA",
  DANCE = "HSSA",
  DANCEW = "HSSA",
  EACS = "HSSA",
  EARTH = "Science",
  EARTHW = "Science",
  ECE = "Engineering",
  ECON = "Economics/Business",
  ED = "HSSA",
  EDS = "Math/Stats",
  EEMB = "Science",
  ENGL = "HSSA",
  ENGR = "Engineering",
  ENV = "Science",
  ES = "HSSA",
  ESS = "HSSA",
  ESM = "Economics/Business",
  FAMST = "HSSA",
  FAMSTW = "HSSA",
  FEMST = "HSSA",
  FEMSTW = "HSSA",
  FLMST = "HSSA",
  FR = "HSSA",
  GEOG = "Science",
  GEOL = "Science",
  GER = "HSSA",
  GLOBL = "HSSA",
  GPS = "HSSA",
  GREEK = "HSSA",
  HEB = "HSSA",
  HIST = "HSSA",
  INT = "HSSA",
  ITAL = "HSSA",
  JAPAN = "HSSA",
  JAPANW = "HSSA",
  KOR = "HSSA",
  LAIS = "HSSA",
  LATIN = "HSSA",
  LING = "HSSA",
  LIT = "HSSA",
  MARSC = "Science",
  MAT = "Science",
  MATH = "Math/Stats",
  MATRL = "Science",
  MCDB = "Science",
  ME = "Engineering",
  MES = "HSSA",
  MS = "HSSA",
  MUS = "HSSA",
  PHIL = "HSSA",
  PHYS = "Science",
  POL = "HSSA",
  PORT = "HSSA",
  PSTAT = "Math/Stats",
  PSTATW = "Math/Stats",
  PSY = "HSSA",
  RG = "HSSA",
  RUSS = "HSSA",
  SHS = "HSSA",
  SLAV = "HSSA",
  SOC = "HSSA",
  SPAN = "HSSA",
  THTR = "HSSA",
  TMP = "Economics/Business",
  "W&L" = "HSSA",
  WRIT = "HSSA"
)

grades_clean <- grades_clean |>
  mutate(dep_group = dept_mapping[dept])
```

```{r}
dept_mapping_rmp <- c( 
  "Accounting" = "Economics/Business", 
  "African-American Studies" = "HSSA", 
  "Anthropology" = "HSSA", 
  "Architecture" = "HSSA", 
  "Art" = "HSSA", 
  "Art History" = "HSSA", 
  "Asian American Studies" = "HSSA", 
  "Biochemistry" = "Science", 
  "Biology" = "Science", 
  "Black Studies" = "HSSA", 
  "Business" = "Economics/Business", 
  "Chemical Engineering" = "Engineering", 
  "Chemistry" = "Science", 
  "Chicano Studies" = "HSSA", 
  "Classics" = "HSSA", 
  "Communication" = "HSSA", 
  "Computer Science" = "Engineering", 
  "East Asian Languages & Cultures" = "HSSA", 
  "Economics" = "Economics/Business", 
  "Education" = "HSSA", 
  "Electrical & Comp. Engineering" = "Engineering", 
  "Engineering" = "Engineering", 
  "English" = "HSSA", 
  "Environmental Science" = "Science", 
  "Environmental Science & Mgmt." = "Science", 
  "Environmental Studies" = "Science", 
  "Ethnic Studies" = "HSSA", 
  "Exercise & Sport Studies" = "HSSA", 
  "Feminist Studies" = "HSSA", 
  "Film" = "HSSA", 
  "Finance" = "Economics/Business", 
  "Fine Arts" = "HSSA", 
  "French" = "HSSA", 
  "French & Italian" = "HSSA", 
  "Geography" = "Science", 
  "Geology" = "Science", 
  "Germanic & Russian Studies" = "HSSA", 
  "Global & International Studies" = "HSSA", 
  "Global Studies" = "HSSA", 
  "Graphic Arts" = "HSSA", 
  "Health Science" = "Science", 
  "History" = "HSSA", 
  "Humanities" = "HSSA", 
  "Information Science" = "Engineering", 
  "International Studies" = "HSSA", 
  "Languages" = "HSSA", 
  "Life Science" = "Science", 
  "Linguistics" = "HSSA", 
  "Literature" = "HSSA", 
  "Mathematics" = "Math/Stats", 
  "Mechanical Engineering" = "Engineering", 
  "Music" = "HSSA", 
  "Philosophy" = "HSSA", 
  "Physical Education" = "HSSA", 
  "Physics" = "Science", 
  "Political Science" = "HSSA", 
  "Psychology" = "HSSA", 
  "Religion" = "HSSA", 
  "Religious Studies" = "HSSA", 
  "Science" = "Science", 
  "Social Science" = "HSSA", 
  "Sociology" = "HSSA", 
  "Spanish" = "HSSA", 
  "Spanish & Portuguese" = "HSSA", 
  "Statistics" = "Math/Stats", 
  "Theater" = "HSSA", 
  "Theology" = "HSSA", 
  "Women's Studies" = "HSSA", 
  "Writing" = "HSSA"
)

rmp_clean <- rmp_clean |>
  mutate(dep_group = dept_mapping_rmp[Department])
```


```{r}
grades_clean <- grades_clean[!is.na(grades_clean$dep_group), ] |>
  filter(nLetterStudents >=5)

rmp_clean <- rmp_clean[!is.na(rmp_clean$dep_group), ] 
```

# Merging Data
```{r, warning=FALSE}
# Standardizing professor names for joining
rmp_named <- rmp_clean |> 
  mutate(nameTransformed = str_to_upper(str_c(word(Name, -1), " ", 
                                              substr(word(Name, 1), 1, 1))))

grades_named <- grades_clean |> 
  mutate(nameTransformed = if_else(str_count(instructor, "\\s+") == 2, 
                                   str_remove(instructor, " \\w+$"), instructor))

# Removing extra dep_group from rmp since it already is there for course grades
rmp_named <- rmp_named |> 
  select(-dep_group)

# Performing inner join to combine datasets
joined_data <- inner_join(rmp_named, grades_named, by = "nameTransformed") |> 
  select(-instructor, -nameTransformed, -Department, -dept, -course, ,-quarter, -year) 

# Aggregating data to be by professor, rather than course
final_data <- joined_data |> 
  group_by(Name) |>
  summarize(
    avgGPA = round(mean(avgGPA), 2),
    rating = mean(`Quality Rating`),
    n_ratings = mean(`Number of Ratings`),
    take_again = first(`Would Take Again`),
    difficulty = first(Difficulty),
    rating_group = first(qualitygroup),
    difficulty_group = first(difficultygroup),
    totalA = sum(totalA), 
    totalB = sum(totalB),
    totalC = sum(totalC),
    totalD = sum(totalD),
    totalF = sum(totalF),
    dep_group = first(dep_group)
  )

final_data
```

## Weighting Average GPA Values
```{r}
gpa_weighted <- grades_clean |>
  group_by(dep_group) |>
  summarise(avgGPA = sum(avgGPA * nLetterStudents, na.rm = TRUE) / sum(nLetterStudents, na.rm = TRUE))

head(gpa_weighted)
```

# Writing Cleaned Data to CSV
```{r}
write_csv(gpa_weighted, "gpa_weighted.csv")
write_csv(final_data, "final_data.csv")
```