---
title: "RateMyProf"
output: html_document
date: "2024-11-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing Dependecies
#hello
```{r, echo = FALSE}
library(tidyverse)
```

Uploading the CSV for Rate My Professor Data and Daily Nexus Data

```{r, echo = FALSE}
rmpdata <- read_csv("/Users/laurenceliao/Downloads/teacher_ratings.csv")
nrow(rmpdata)
```
```{r}
coursegradesdata <- read_csv("/Users/laurenceliao/Downloads/courseGrades.csv")
nrow(coursegradesdata)
```


# Data Cleaning

- Changing the ratings column to remove the "ratings" at the end
- filter to only have those with ratings
- remove professor duplicates


```{r, echo = FALSE}
rmpdata <- rmpdata |> mutate(`Number of Ratings` = as.numeric(gsub("[^0-9]", "", `Number of Ratings`))) |> filter(`Number of Ratings` > 5)
nrow(rmpdata)
rmpdata <- rmpdata |> distinct(Name, .keep_all = TRUE)
nrow(rmpdata)
```

There were a total of 291 professors that had 0 ratings and 20 duplicate ratings. The number of professors we are now analyzing is 2804.

- Group grades so that each category is just A, B, C, D, F regardless of plus or minus

```{r}
coursegradesdata <- coursegradesdata |> mutate(totalA = A + Ap + Am,
                                               totalB = B + Bp + Bm,
                                               totalC = C + Cp + Cm, 
                                               totalD = D + Dp + Dm,
                                               totalF = F)
```

- Group quality ratings to 0-1 group, 1-2, etc.

```{r}
rmpdata <- rmpdata |> mutate(qualitygroup = case_when(`Quality Rating` >= 0 & `Quality Rating` < 1 ~ "0-1",
                                                      `Quality Rating` >= 1 & `Quality Rating` < 2 ~ "1-2",
                                                     `Quality Rating` >= 2 & `Quality Rating` < 3 ~ "2-3",
                                                      `Quality Rating` >= 3 & `Quality Rating` < 4 ~ "3-4",
                                                      `Quality Rating` >= 4 & `Quality Rating` <= 5 ~ "4-5"),
                             difficultygroup = case_when(Difficulty >= 0 & Difficulty < 1 ~ "0-1",
                                                      Difficulty >= 1 & Difficulty < 2 ~ "1-2",
                                                      Difficulty >= 2 & Difficulty < 3 ~ "2-3",
                                                      Difficulty >= 3 & Difficulty < 4 ~ "3-4",
                                                      Difficulty >= 4 & Difficulty <= 5 ~ "4-5"))
```

- Merge datasets using name of professor

```{r}
rmpdatatest <- rmpdata |> mutate(nameTransformed = str_to_upper(str_c(word(Name, -1), " ", substr(word(Name, 1), 1, 1))))
coursegradesdatatest <- coursegradesdata |> 
  mutate(nameTransformed = if_else(str_count(instructor, "\\s+") == 2, str_remove(instructor, " \\w+$"), instructor))

finaldatatest <- inner_join(rmpdatatest, coursegradesdatatest, by = "nameTransformed") |> select(-instructor, -nameTransformed, -Department)
```

```{r}
econdata <- finaldatatest |> filter(dept == "ECON" & nLetterStudents > 0)
grouped_econdata <- econdata |> group_by(Name) |> summarize(A = sum(totalA),
                                         B = sum(totalB), 
                                         C = sum(totalC),
                                         D = sum(totalD),
                                         F = sum(totalF)) |>
  ungroup() |> mutate(totalGrades = A + B + C + D + F)


econdatalist <- econdata |> distinct(Name, .keep_all = TRUE) |> arrange(Name) |> select(Name, qualitygroup, difficultygroup)

merged_econ <- inner_join(grouped_econdata, econdatalist, by = "Name")

```


# Data Analysis + Visualization

- Aggregate grades based department and qualitygroup or difficultygroup

```{r}
grouped <- merged_econ |> group_by(qualitygroup) |> summarize(totalA = sum(A),
                                                              totalB = sum(B),
                                                              totalC = sum(C),
                                                              totalD = sum(D),
                                                              totalF = sum(F),
                                                              total = sum(totalGrades),
                                                              countOfProfs = n())

grouped
```











